When Precision Therapy Creates Inprecise Choices

### 2022-12-16 Manuscript.qmd
revision

The third key criteria supporting the use of prophylaxis is diagnostic difficulty. The availability of improved diagnostic tools (serum galactomannan, PCR) and high-resolution CT scanning has improved the early diagnosis of these infections allowing for possible or a preemeptive treatment approaches- i.e. initiation of antifungal therapy after detection of early fungal biomarkers and suggestive chest CT findings. Preemptive management approaches for IFD have been shown to be as effective as empirical treatment in high-risk patients with persistent fever [@morrissey2013]. An EORTC-lead randomized trial demonstrated that preemptive antifungal strategy was safe for high-risk neutropenic patients given fluconazole as prophylaxis, halving the number of patients receiving mould-active antifungals without excess mortality or IFDs [@Maertens2022]. However, primary antifungal prophylaxis with posaconazole has not been directly compared with empiric or preemptive therapy in a randomized clinical trial during RIC for AML.

@@

Non-triazole based agents, namely echinocandins or liposomal amphotericin B, are often considered as alternatives to triazoles in patients who require SMKIs or are enrolled in clinical trials with new targeted therapies that preclude use triazoles because of QT-prolongation risk [@bruggemann_etal22]. Currently, only micafungin is approved for prophylaxis of Candida infection in patients undergoing allogeneic hematopoietic stem cell transplantation or patients who are expected to have neutropenia (absolute neutrophil count \< 500 cells / µl) for 10 or more days. Current guidelines for the treatment of invasive aspergillosis do not recommend echinocandin monotherapy as a frontline treatment for invasive aspergillosis due to limited evidence of efficacy in patients with probable or proven IA [@ullmann2012a; @viscoli2009; @herbrecht2010a]. A recent open-label randomized trial comparing caspofungin to fluconazole prophylaxis in children and adolescents undergoing treatment for AML found significantly lower cumulative incidence of proven or probable aspergillosis in patient who received caspofungin versus fluconazole (0.5% vs. 3%, P=0.04) [@fisher_etal19a], although the rates of IA in the fluconazole arm were lower than reported in previous adult studies that compared posaconazole suspension to fluconazole/itraconazole (1% vs. 7%, P=0.04) [@cornely2007J]. Some [@gomes2014] but not all [@Epstein2018a];El Cheikh et al., 2019, #101153} single-center observational studies examining echinocandin prophylaxis during RIC for AML have reported higher rates of breakthrough IFD in patients receiving echinocandin prophylaxis versus triazoles. Therefore, it is still unclear whether echinocandin prophylaxis, despite the improved safety profile when administered with SMKIs, is as effective as posaconazole prophylaxis for preventing invasive mold disease.

Echinocandin prophylaxis may be associated with higher risk for breakthrough fungal infections caused by rare yeast {Stanzani et al., 2019, #287617;Gomes et al., 2013, #95041; El Zein et al., 2020, #229138} and intrinsically-resistant mold such as *Fusarium* spp. and Mucorales compared to mould-active triazoles {Gomes et al., 2013, #95041}. In centers or regions where these infections are more common, careful consideration should be given to early diagnosis of these breakthrough infections given their propensity for fulminant infections and rapid dissemination.

Intravenous liposomal amphotericin B (L-AMB) is occasionally used off-label as a prophylaxis agent, although the optimal dosing approach and efficacy is not well established. This approach was often considered adult patients with acute lymphoblastic leukemic (ALL), who frequently receive chemotherapy regimens containing vinca alkaloids, which are metabolized through CYP3A4 {Stafylidis et al., 2022, #235425}. Cornely et al. performed a double-blind multicenter Phase 3 study comparing twice-weekly prophylactic liposomal amphotericin B (L-AMB) at 5 mg/kg intravenously or placebo twice during remission-induction treatment {Cornely et al., 2017, #34343}. Rates of proven and probable IFD were similar between L-AMB and placebo (7.9% vs. 11.7) with similar mortality rates. Hypokalaemia (35.0% vs. 17.8%) and increased creatinine (\> 1.5x baseline, 3.0% vs. 0%) were significantly more frequent in patients who received L-AMB prophylaxis. It is unclear if the limited efficacy was a reflection of the infrequent dosing regimen- therefore, if L-AMB is considered it may be more judicious to administer a daily 3 mg/kg dose for a define period (i.e. \< 21 days) if in unique cases L-AMB prophylaxis is used in patients who cannot tolerate triazoles or echinocandins.

Both echinocandins and liposomal amphotericin B must be administered intravenously, making them more problematic to use in the outpatient setting in patients taking SMKIs as part of a consolidation- maintenance regimen. An oral echinocandin-like triterpenoid recently approved for the treatment of vuvlvovaginal candidiasis, ibrexafungerp, is currently under investigations

If a triazole must be given, is isavuconazole the preferred?

What will be the potential role of new antifungals?

Table

Should the dose of the SMKI by empirically reduced to account for the effect of triazole CYP3A4/5 inhibition?

Venetoclax

FLT3 inhibitors

IDH1/2 inhibitors

What is the role of therapeutic drug monitoring?

>>>>

The third key criteria supporting the use of prophylaxis is diagnostic difficulty. The availability of improved diagnostic tools (serum galactomannan, PCR) and high-resolution CT scanning has improved the early diagnosis of these infections allowing for possible or a preemeptive treatment approaches- i.e. initiation of antifungal therapy after detection of early fungal biomarkers and suggestive chest CT findings. Preemptive management approaches for IFD have been shown to be as effective as empirical treatment in high-risk patients with persistent fever [@morrissey_etal13]. An EORTC-lead randomized trial demonstrated that preemptive antifungal strategy was safe for high-risk neutropenic patients given fluconazole as prophylaxis, halving the number of patients receiving mould-active antifungals without excess mortality or IFDs [@maertens_etal22]. However, primary antifungal prophylaxis with posaconazole has not been directly compared with empiric or preemptive therapy in a randomized clinical trial during RIC for AML.

@@

Non-triazole based agents, namely echinocandins or liposomal amphotericin B, are often considered as alternatives to triazoles in patients who require SMKIs or are enrolled in clinical trials with new targeted therapies that preclude use triazoles because of QT-prolongation risk [@bruggemann_etal22]. Currently, only micafungin is approved for prophylaxis of Candida infection in patients undergoing allogeneic hematopoietic stem cell transplantation or patients who are expected to have neutropenia (absolute neutrophil count \< 500 cells / µl) for 10 or more days. Current guidelines for the treatment of invasive aspergillosis do not recommend echinocandin monotherapy as a frontline treatment for invasive aspergillosis due to limited evidence of efficacy in patients with probable or proven IA [@Ullmann2018; @Viscoli2009; @Herbrecht2010]. A recent open-label randomized trial comparing caspofungin to fluconazole prophylaxis in children and adolescents undergoing treatment for AML found significantly lower cumulative incidence of proven or probable aspergillosis in patient who received caspofungin versus fluconazole (0.5% vs. 3%, P=0.04) [@Fisher2019], although the rates of IA in the fluconazole arm were lower than reported in previous adult studies that compared posaconazole suspension to fluconazole/itraconazole (1% vs. 7%, P=0.04) [@cornely2007J]. Some [@gomes2014] but not all [@Epstein2018;@El-Cheikh2013]single-center observational studies examining echinocandin prophylaxis during RIC for AML have reported higher rates of breakthrough IFD in patients receiving echinocandin prophylaxis versus triazoles. Therefore, it is still unclear whether echinocandin prophylaxis, despite the improved safety profile when administered with SMKIs, is as effective as posaconazole prophylaxis for preventing invasive mold disease.

Echinocandin prophylaxis may be associated with higher risk for breakthrough fungal infections caused by rare yeast [@stanzani_etal19;@Gomes2014;@El2020]and intrinsically-resistant mold such as Fusarium spp. and Mucorales compared to mould-active triazoles [@Gomes2014]. In centers or regions where these infections are more common, careful consideration should be given to early diagnosis of these breakthrough infections given their propensity for fulminant infections and rapid dissemination.

Intravenous liposomal amphotericin B (L-AMB) is occasionally used off-label as a prophylaxis agent, although the optimal dosing approach and efficacy is not well established. This approach was often considered adult patients with acute lymphoblastic leukemic (ALL), who frequently receive chemotherapy regimens containing vinca alkaloids, which are metabolised through CYP3A4 [@Stafylidis2022]. Cornely et al. performed a double-blind multicenter Phase 3 study comparing twice-weekly prophylactic liposomal amphotericin B (L-AMB) at 5 mg/kg intravenously or placebo twice during remission-induction treatment [@Cornely2017]. Rates of proven and probable IFD were similar between L-AMB and placebo (7.9% vs. 11.7) with similar mortality rates. Hypokalaemia (35.0% vs. 17.8%) and increased creatinine (\> 1.5x baseline, 3.0% vs. 0%) were significantly more frequent in patients who received L-AMB prophylaxis. It is unclear if the limited efficacy was a reflection of the infrequent dosing regimen- therefore, if L-AMB is considered it may be more judicious to administer a daily 3 mg/kg dose for a define period (i.e. \< 21 days) if in unique cases L-AMB prophylaxis is used in patients who cannot tolerate triazoles or echinocandins.

Both echinocandins and liposomal amphotericin B must be administered intravenously, making them more problematic to use in the outpatient setting in patients taking SMKIs as part of a consolidation- maintenance regimen. An oral echinocandin-like triterpenoid recently approved for the treatment of vuvlvovaginal candidiasis, ibrexafungerp, is currently under investigation for the treatment of invasive candidiasis and treatment of invasive aspergillosis in combination with voriconazole[@johnson2021].

##

# Should the dose of the SMKI by empirically reduced to account for the effect of triazole CYP3A4/5 inhibition?

## Venetoclax

## FLT3 inhibitors

## IDH1/2 inhibitors

## What is the role of therapeutic drug monitoring?

# What are the risks of continuing triazoles and SMKIs at full-dose?
